\documentclass[11pt]{article}

\renewcommand\familydefault{\sfdefault}
 
%input preamble and macros
\input{../../slides/macros/preamble}
\input{../../slides/macros/macros}

 
\usepackage[margin=1in]{geometry} 
\usepackage{amsmath,amsthm,amssymb}
 
%\newcommand{\N}{\mathbb{N}}
%\newcommand{\Z}{\mathbb{Z}}

% Named environments (no counters) 
\newenvironment{theorem}[2][Theorem]{\begin{trivlist}
\item[\hskip \labelsep {\bfseries #1}\hskip \labelsep {\bfseries #2.}]}{\end{trivlist}}
\newenvironment{lemma}[2][Lemma]{\begin{trivlist}
\item[\hskip \labelsep {\bfseries #1}\hskip \labelsep {\bfseries #2.}]}{\end{trivlist}}
%\newenvironment{exercise}[2][Exercise]{\begin{trivlist}
%\item[\hskip \labelsep {\bfseries #1}\hskip \labelsep {\bfseries #2.}]}{\end{trivlist}}
\newenvironment{problem}[2][Problem]{\begin{trivlist}
\item[\hskip \labelsep {\bfseries #1}\hskip \labelsep {\bfseries #2.}]}{\end{trivlist}}
\newenvironment{question}[2][Question]{\begin{trivlist}
\item[\hskip \labelsep {\bfseries #1}\hskip \labelsep {\bfseries #2.}]}{\end{trivlist}}
\newenvironment{corollary}[2][Corollary]{\begin{trivlist}
\item[\hskip \labelsep {\bfseries #1}\hskip \labelsep {\bfseries #2.}]}{\end{trivlist}}
 
% Environments with counters
\newtheoremstyle{myplain} {3mm}% ⟨Space above⟩
{3mm}% ⟨Space below⟩
{}% ⟨Body font⟩
{}% ⟨Indent amount⟩
{\bfseries}% ⟨Theorem head font⟩
{.}% ⟨Punctuation after theorem head⟩
{.5em}% ⟨Space after theorem head⟩2
{}% ⟨Theorem head spec (can be left empty, meaning ‘normal’)⟩

\theoremstyle{myplain}
\newtheorem{exercise}{Exercise}

\theoremstyle{definition} % no italics
\newtheorem{subexercise}{}[exercise]

\newcommand{\subex}[1]{\begin{subexercise}#1\end{subexercise}}

\begin{document}
 
% --------------------------------------------------------------
%                         Start here
% --------------------------------------------------------------
 
\title{Assignment: Vending machines in mCRL2}
\author{Jos\'{e} Proen\c{c}a\\
Arquitectura e C\'alculo -- 2015/2016} 
 
\maketitle
 
\section*{Modelling a vending machine}

%-----------------------------------------------------------------------------
\begin{exercise} \label{ex:vm1}
We specify a very primitive vending machine and a user in mCRL2 below. 
After inserting a coin of 50 cents, the user can push the button for an apple.

\begin{lstlisting}
act
  ins50, optA, acc50, putA, coin, ready ;
proc
  User = ins50 . optA . User ;
  Mach = acc50 . putA . Mach ;
init
  allow(
    { coin, ready },
    comm(
      { ins50|acc50 -> coin, optA|putA -> ready },
      User || Mach
  ) ) ;  
\end{lstlisting}  

The specification is split into three sections:
\code{act}, a declaration of actions of 6 actions,
\code{proc}, the definition of 2 processes, and
\code{init}, the initialisation of the system.


\subex{
Produce the labelled transition system (LTS) of this specification using (1) \bash{mcrl22lps} to linearise the system and (2) \bash{lps2lts} to produce the final LTS.}

\subex{Visualise the previous LTS with \bash{ltsgraph}.
\textbf{Show a screenshot of the LTS (after rearranging as needed.}}

\subex{Specify and visualise the LTS of a similar specification, obtained by omitting the restrictions \code{allow} and \code{comm}.
\textbf{How many states do you expect (and count), and why?}}

\end{exercise}

%----------------------------------------------
\begin{exercise} \label{ex:vm2}
Next, we add a chocolate bar to the assortment of the vending machine. A chocolate bar costs 1 euro, an apple 50 cents. The machine will now accept coins of 50 cents and 1 euro. The scenarios allowed are (i) insertion of 50 cents and purchasing an apple, (ii) insertion of 50 cent twice or 1 euro once and purchasing a chocolate bar. Additionally, after insertion of money, the user can push the change button, after which the inserted money is returned.

\subex{Extend the following mCRL2 specification to describe the vending machine sketched above, and \textbf{save the resulting specification as \bash{vm2.mcrl2}}. The actions that are involved, and a possible specification of the \code{Mach} process have been given. The machine is required to perform a \code{prod} action for administration purposes.}

\begin{lstlisting}
act
  ins50, ins100, acc50, acc100, coin50, coin100, ret50, ret100 ;
  optA, optC, chg50, chg100, putA, putC, prod, 
  readyA, readyC, out50, out100 ;

proc
  User = 
    %% 1 %%

  Mach = 
    acc50.( putA.prod + acc50.( putC.prod + ret100 ) + ret50 ).Mach +
    acc100.( putA.prod.ret50 + putC.prod + ret100 ).Mach ;

init
  %% 2 %%
\end{lstlisting}

\subex{Linearise your specification, build the LTS, and visualise the result LTS using \bash{mcrl22lps}, \bash{lps2lts}, and \bash{ltsgraph}.
\textbf{Show a screenshot of the obtained graph.}}

\end{exercise}


% --------------------------------------------------------------
\begin{exercise}
The same vending machine is now specified below using data parameters to capture different coins and products.

\begin{lstlisting}
sort Coin = struct c50ct | c1eur;
     Product = struct Apple | Chocolate;

act
  ins, acc, coin, ret, chg, out: Coin;
  opt, put, ready: Product;
  prod;

proc
  User = 
    ins(c50ct).( opt(Apple) + ins(c50ct).( opt(Chocolate) + chg(c1eur) )
               + chg(c50ct) ).User +
    ins(c1eur).( opt(Apple).chg(c50ct) + opt(Chocolate) + chg(c1eur) ).User ;

  Mach = 
    acc(c50ct).( put(Apple).prod + acc(c50ct).( put(Chocolate).prod + ret(c1eur) )
               + ret(c50ct) ).Mach +
    acc(c1eur).( put(Apple).prod.ret(c50ct) + put(Chocolate).prod + ret(c1eur) ).Mach ;

init
  allow(
    { coin, ready, out, prod },
    comm(
      { ins|acc -> coin,
        chg|ret -> out,
        opt|put -> ready },
      User || Mach
  ) ) ;  
\end{lstlisting}

\subex{Modify this specification such that all coins of denomination 2eur, 1eur, 50ct, and 20ct can be inserted. The machine accumulates upto a total of 2.50 eur. If sufficient credit, an apple or chocolate bar is supplied after selection. Money is returned after pressing the change button.
\textbf{Show this updated specification.}}

\subex{\bfseries Visualise and include a screenshot of the new machine.}

\end{exercise}

% --------------------------------------------------------------
\section*{LTS Equivalence}


\begin{exercise}  
Recall the \bash{vm2.mcrl2} specification from Exercise~\ref{ex:vm2} of a system performing \code{coin50} and \code{coin100} actions as well as so-called $\tau$-steps.

\subex{Build a specification \bash{vm2-taus.mcrl2} that is the same as \bash{vm2.mcrl2} after hiding the actions \code{readyA}, \code{readyC}, \code{out50}, \code{out100}, \code{prod}. \textbf{Show this specification.}}

\subex{Using the \bash{ltscompare} tool, compare your model under branching bisimilarity with the LTS of the system \bash{vm2-taus} with the LTS of the system \bash{vm2} after hiding the actions \code{readyA}, \code{readyC}, \code{out50}, \code{out100}, \code{prod}. For that use the following command.}

\begin{lstlisting}[style=bash]
$\texttt{\$}$ ltscompare --equivalence=branching-bisim \
             --tau=out50,out100,readyA,readyC,prod vm2.lts vm2-taus.lts
\end{lstlisting}


\subex{Using \bash{ltsconvert}, minimise the LTS for \bash{vm2.mcrl2} with respect to branching bisimulation after hiding the same actions as before (using the \code{hide} operation in mcrl2).
\textbf{Visualise and include a screenshot of the minimised LTS.}
Compare the minimised LTSs and \bash{vm2-taus.lts} visually using \bash{ltsgraph}.}

%\begin{lstlisting}[style=bash]
%$\texttt{\$}$ ltsconvert -ebranching-bisim --tau=out10,out20,readyA,readyC,prod vm02.svc vm02min.svc}

\end{exercise}


% --------------------------------------------------------------
\section*{Verification of the vending machines}

\begin{exercise}
Recall the simple LTS from Exercise~\ref{ex:vm1}.

\subex{\label{ex:ver1}
What does the property ``\code{[true*]<ready>true}'' mean? Does it hold?}

\subex{\label{ex:ver2}
What does the property ``\code{[true*.ready.!coin]false}'' mean? Does it hold? Note: \code{!coin} represents the complement of \code{coin}, i.e., any action that is not \code{coin}.}

\subex{\label{ex:ver3}
Write a property that expresses that, at any moment, it is possible to reach a state where a coin can be inserted.
}

\subex{
Use the mCRL2 toolset to verify if the 3 properties above hold (from Exercises~\ref{ex:ver1},~\ref{ex:ver2}, and~\ref{ex:ver3}).
This involves 2 steps for each property:
\begin{enumerate}
\item use ``\bash{lps2pbes <lps-file> -f <file-with-formula> <output.pbes>}'' to produce a system of boolean equations, and
\item use ``\bash{pbes2bool <output.pbes>}'' to evaluate if these equations have a solution.  
\end{enumerate}
}

\subex{Adapt the 3 properties (if needed) to the vending machine specified in Exercise~\ref{ex:vm2} and verify them using the mCRL2 toolset. \textbf{Show these properties and the result of the verification.}}

\end{exercise}
 
% --------------------------------------------------------------
%     You don't have to mess with anything below this line.
% --------------------------------------------------------------
 
\end{document}